---
title: "Size cline analysis"
author: "Samuel Perini"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bbmle)
library(magrittr)
library(dplyr)
library(pander)
```


```{r cline function 2 centres 4 sd, message=FALSE}
CZ_all = read.csv("../../data/CZ_all_mating_clean.csv", sep = ";")

cline_2c4s <- function(phen,position,sex,cl,cr,lwl,lwr,crab,wave,zs_c,zs_w,sc,shl,sh,sw){
  wl = exp(lwl)
  wr = exp(lwr)
  # sc = exp(lsc)
  # sh = exp(lsh)
  # sw = exp(lsw)
  # left cline
  p_xl <- 1-1/(1+exp(0-4*(position-cl)/wl))  # decreasing
  z_xl <- crab+(wave-crab)*p_xl  # z_xl is expected phenotype for left cline
  z_xl[sex=="female"] <- z_xl[sex=="female"] + zs_c + (zs_w-zs_c)*p_xl[sex=="female"]
  s_xl <- sqrt(sc^2 + 4*p_xl*(1-p_xl)*shl^2 + (p_xl^2)*(sw^2-sc^2))

  # right cline
  p_x <- 1/(1+exp(0-4*(position-cr)/wr))  # increasing
  z_x <- crab+(wave-crab)*p_x  # z_x is expected phenotype for the right cline
  z_x[sex=="female"] <- z_x[sex=="female"] + zs_c + (zs_w-zs_c)*p_x[sex=="female"]
  s_x <- sqrt(sc^2 + 4*p_x*(1-p_x)*sh^2 + (p_x^2)*(sw^2-sc^2))

  # combined cline
  z_x[z_x < z_xl] <- z_xl[z_x < z_xl]
  s_x[z_x < z_xl] <- s_xl[z_x < z_xl]
  minusll <- -sum(dnorm(phen,z_x,s_x,log=T))
  if(crab > wave){minusll <- minusll+1000}
  if(cl > cr){minusll <- minusll+1000}
  return(minusll)
}

p = "CZB"
theta.init = list(cl=70,cr=150,lwl=1.6,lwr=3.9,crab=-2.5,wave=-1.5,zs_c=-0.1,zs_w=-0.1,sc=0.2,shl=0.3,sh=0.3,sw=0.2)
mle.cline.2c3s = mle2(cline_2c3s, theta.init,
                      control=list(parscale=abs(unlist(theta.init))),
                      data=list(phen=-log(CZ_all$length_mm[CZ_all$shore==p]),
                                position=CZ_all$DistAlongPath[CZ_all$shore==p],
                                sex=CZ_all$sex[CZ_all$shore==p]))

(cline_sum = summary(mle.cline.2c3s))
AIC(mle.cline.2c3s)
```

**Note**:

* I have tried to change the left cline as increasing and the right cline as decreasing but I have not succeded. Perhaps I can have a look at your cline script and change mine accordingly.

* _zs\_c_ and _zs\_w_ are not properly estimated because their values are the same as the intial values. Can you see anything I can change to improve the fitting? 

\

Back-transformation of the cline paramters:
```{r transpars}
cline_params = coef(cline_sum)[,1]
cline_params[(5:8)] = abs(cline_params[(5:8)])
cline_params[(-1:-2)] = exp(cline_params[(-1:-2)])
names(cline_params) = c("cl","cr","wl","wr","crab_mm","wave_mm","zs_c_mm","zs_w_mm","sc","sh","sw")
p_xl = 0.5
cline_params['sh'] = sqrt(cline_params['sc']^2 + 4*p_xl*(1-p_xl)*cline_params['sh']^2 + (p_xl^2)*(cline_params['sw']^2-cline_params['sc']^2))
pander::pander(x = cline_params %>% round(2))
```


```{r fig size vs distance, fig.pos='h', echo=FALSE}
par(mar = c(4, 4, 0.5, 0.5))
plot(CZ_all$DistAlongPath[CZ_all$shore==p], CZ_all$length_mm[CZ_all$shore==p], pch=19,
     xlab = "Distance (m)", ylab = "Size", main = "CZB")
abline(v = c(cline_params['cl'], cline_params['cr']), col=c('blue', 'blue'))
```

